### 2023-08-10 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

---
title: 'MSCA 31006: Final Project - AQI Forecasting'
author: "Aashish Singh"
date: "8/10/2023"
output: html_document
---

```{r, tidy=TRUE}
#install.packages(c("forecast", "expsmooth", "seasonal")) 
library(TTR)
library(forecast) 
library(tseries) 
library(expsmooth) 
library(fpp2)
library(seasonal)
library(MASS)
library(TSA)
library(forecast)
library(ggplot2)
library(tseries)
library(imputeTS)
library(vars)
library(timetk)
library(Metrics)
library(lmtest)
library(lubridate)
library(xts)
library(dplyr)
```

#### Step 1: Load data and plot the time series

```{r, tidy=TRUE}
path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
fresno_pm2_5 <- read.csv(paste0(path, "fresno_pm2_5_pollutant_data.csv"), 
                         na.strings=c("", "NA"))
fresno_ozone <- read.csv(paste0(path, "fresno_ozone_pollutant_data.csv"), 
                         na.strings=c("", "NA"))

# Convert data into ts objects
fresno_pm2_5_ts <- ts(fresno_pm2_5$PM2.5.AQI.Value, start = c(1999,1,1), frequency = 365.25)
fresno_ozone_ts <- ts(fresno_ozone$Ozone.AQI.Value, start = c(1999,1,1), frequency = 365.25)

plot(fresno_pm2_5_ts, main="Fresno PM 2.5 AQI")
plot(fresno_ozone_ts, main="Fresno Ozone AQI")
```

#### Step 2: Check if data is stationary (KPSS/ADF Test)

```{r, tidy=TRUE}

kpss.test(fresno_pm2_5_ts)
# The reported p-value is 0.01, which is smaller than 0.05, and would suggest 
# that we reject the null hypothesis of level stationarity and conclude that 
# there is evidence that the data is non-stationary

adf.test(fresno_pm2_5_ts)
# Similarly the Augumented Dickey-Fuller test results in p-value of 0.01 (<0.05) 
# where we do reject the null hypothesis that the time series is non-stationary
# and thus data is stationary.

# These are opposite results so we use ACF PACF plots to see if the series is stationary
acf(fresno_pm2_5_ts, main = "ACF of Original Time Series")
pacf(fresno_pm2_5_ts, main = "PACF of Original Time Series")

```

#### Step 3: Extract weekly values for time series and plot the series

```{r, tidy=TRUE}
# Create weekly data
fresno_pm2_5_df <- as.data.frame(fresno_pm2_5_ts)
fresno_pm2_5_df$Date <- mdy(fresno_pm2_5$Date)
colnames(fresno_pm2_5_df) <- c("PM2.5.AQI.Value", "Date")

fresno_pm2_5_xts <- as.xts(fresno_pm2_5_df, order.by=fresno_pm2_5_df$Date)
fresno_pm2_5_xts <- fresno_pm2_5_xts[,-2]
fresno_pm2_5_weekly <- apply.weekly(fresno_pm2_5_xts, mean)

fresno_pm2_5_weekly_ts <- ts(fresno_pm2_5_weekly["19990103/20230805"], start = c(1999,1),frequency = 52.18)
plot(fresno_pm2_5_weekly_ts)

kpss.test(fresno_pm2_5_weekly_ts)

adf.test(fresno_pm2_5_weekly_ts)

shapiro.test(fresno_pm2_5_weekly_ts)

lambda <- BoxCox.lambda(fresno_pm2_5_weekly_ts)

ts.plot(BoxCox(fresno_pm2_5_weekly_ts, lambda), main="Box-Cox Transformed PM 2.5 AQI")

fresno_pm2_5_weekly_ts_transformed <- BoxCox(fresno_pm2_5_weekly_ts, lambda)

kpss.test(fresno_pm2_5_weekly_ts_transformed)

adf.test(fresno_pm2_5_weekly_ts_transformed)

acf(fresno_pm2_5_weekly_ts_transformed, main = "ACF of Box-Cox Transformed Time Series")
pacf(fresno_pm2_5_weekly_ts_transformed, main = "PACF of Box-Cox Transformed Time Series")

```

#### Step 4: Extract monthly values for time series and plot the series

```{r, tidy=TRUE}
# Create monthly data
fresno_pm2_5_monthly <- apply.monthly(fresno_pm2_5_xts, mean)

fresno_pm2_5_monthly_ts <- ts(fresno_pm2_5_monthly["19990130/20230805"], start = c(1999,1), frequency = 12)
plot(fresno_pm2_5_monthly_ts)

kpss.test(fresno_pm2_5_monthly_ts)

adf.test(fresno_pm2_5_monthly_ts)

acf(fresno_pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
pacf(fresno_pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
```

### 2023-08-11 MSCA_31006_Time_Series_Final_Project.Rmd
Update MSCA_31006_Time_Series_Final_Project.Rmd

@@
 title: 'MSCA 31006: Final Project - AQI Forecasting'
-author: "Aashish Singh"
+author: "Aashish Singh, Jacob Brewer, Nikitha Gopal, Sydney Peters"
 date: "8/10/2023"
@@
 
-shapiro.test(fresno_pm2_5_weekly_ts)
-
-lambda <- BoxCox.lambda(fresno_pm2_5_weekly_ts)
-
-ts.plot(BoxCox(fresno_pm2_5_weekly_ts, lambda), main="Box-Cox Transformed PM 2.5 AQI")
-
-fresno_pm2_5_weekly_ts_transformed <- BoxCox(fresno_pm2_5_weekly_ts, lambda)
-
-kpss.test(fresno_pm2_5_weekly_ts_transformed)
-
-adf.test(fresno_pm2_5_weekly_ts_transformed)
-
-acf(fresno_pm2_5_weekly_ts_transformed, main = "ACF of Box-Cox Transformed Time Series")
-pacf(fresno_pm2_5_weekly_ts_transformed, main = "PACF of Box-Cox Transformed Time Series")
-
 ```
@@
 ```
+
+#### Step 5: Look at ACF / PACF plots
+
+```{r, tidy=TRUE}
+acf(fresno_pm2_5_weekly_ts, main = "ACF of  Time Series")
+pacf(fresno_pm2_5_weekly_ts, main = "PACF of Time Series")
+
+lambda <- BoxCox.lambda(fresno_pm2_5_weekly_ts)
+
+ts.plot(BoxCox(fresno_pm2_5_weekly_ts, lambda), main="Box-Cox Transformed PM 2.5 AQI")
+
+fresno_pm2_5_weekly_ts_transformed <- BoxCox(fresno_pm2_5_weekly_ts, lambda)
+
+kpss.test(fresno_pm2_5_weekly_ts_transformed)
+
+adf.test(fresno_pm2_5_weekly_ts_transformed)
+
+acf(fresno_pm2_5_weekly_ts_transformed, main = "ACF of Box-Cox Transformed Time Series")
+pacf(fresno_pm2_5_weekly_ts_transformed, main = "PACF of Box-Cox Transformed Time Series")
+
+plot(decompose(fresno_pm2_5_weekly_ts_transformed,type="additive"))
+
+```
+
+#### Step 5: Understand the composition of the time series data
+```{r, tidy=TRUE}
+
+#differencing data
+# this can help stablise mean of time series by removing chnges in the level of a time series
+
+fresno_pm2_5_weekly_ts_diff<- diff(fresno_pm2_5_weekly_ts)
+plot(fresno_pm2_5_weekly_ts_diff)
+
+acf(fresno_pm2_5_weekly_ts_diff, main = "ACF of differenced Time Series")
+pacf(fresno_pm2_5_weekly_ts_diff, main = "PACF of differenced Time Series")
+
+decomp <- decompose(fresno_pm2_5_weekly_ts, type="additive")
+plot(decomp)
+
+```

### 2023-08-13 train_test_split.R
Add files via upload

@@ -1,0 +1,21 @@
+LA_pm2_5_ts <- ts(pm2.5_LA$PM2.5.AQI.Value, start = c(1999,3,1), frequency = 365.25)
+
+LA_pm2_5_df <- as.data.frame(LA_pm2_5_ts)
+LA_pm2_5_df$Date <- mdy(pm2.5_LA$Date)
+colnames(LA_pm2_5_df) <- c("PM2.5.AQI.Value", "Date")
+
+LA_pm2_5_xts <- as.xts(LA_pm2_5_df, order.by=LA_pm2_5_df$Date)
+LA_pm2_5_xts <- LA_pm2_5_xts[,-2]
+LA_pm2_5_weekly <- apply.weekly(LA_pm2_5_xts, mean)
+
+LA_pm2_5_weekly_ts <- ts(LA_pm2_5_weekly["19990103/20230805"], start = c(1999,1),frequency = 52.18)
+plot(LA_pm2_5_weekly_ts)
+
+# Create the training dataset: from start to 2020
+train <- window(LA_pm2_5_weekly_ts, start = c(1999, 1), end = c(2019, 52))  # assuming 52 weeks in 2020
+
+# Create the test dataset: from the start of 2020 onwards
+test <- window(LA_pm2_5_weekly_ts, start = c(2020, 1))
+
+plot(train, main = "Training Data")
+plot(test, main = "Testing Data")

### 2023-08-13 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

--- 2023-08-10 MSCA_31006_Time_Series_Final_Project.Rmd
@@
 library(MASS)
+library(stats)
 library(TSA)
@@
 library(dplyr)
+library(tidyr)
+library(sqldf)
 ```
@@
 path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
-fresno_pm2_5 <- read.csv(paste0(path, "fresno_pm2_5_pollutant_data.csv"), 
+pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
                      na.strings=c("", "NA"))
-fresno_ozone <- read.csv(paste0(path, "fresno_ozone_pollutant_data.csv"), 
+ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
                      na.strings=c("", "NA"))
@@
 # Convert data into ts objects
-fresno_pm2_5_ts <- ts(fresno_pm2_5$PM2.5.AQI.Value, start = c(1999,1,1), frequency = 365.25)
-fresno_ozone_ts <- ts(fresno_ozone$Ozone.AQI.Value, start = c(1999,1,1), frequency = 365.25)
+pm2_5_ts <- ts(pm2_5_data$PM2.5.AQI.Value, start = c(1999,1,1), frequency = 365.25)
+ozone_ts <- ts(ozone_data$Ozone.AQI.Value, start = c(1999,1,1), frequency = 365.25)
 
-plot(fresno_pm2_5_ts, main="Fresno PM 2.5 AQI")
-plot(fresno_ozone_ts, main="Fresno Ozone AQI")
+plot(pm2_5_ts, main="Los Angeles PM 2.5 AQI")
+plot(ozone_ts, main="Los Angeles Ozone AQI")
 ```
@@
 
-kpss.test(fresno_pm2_5_ts)
+kpss.test(pm2_5_ts)
 # The reported p-value is 0.01, which is smaller than 0.05, and would suggest 
@@
 
-adf.test(fresno_pm2_5_ts)
+adf.test(pm2_5_ts)
 # Similarly the Augumented Dickey-Fuller test results in p-value of 0.01 (<0.05) 
@@
 # These are opposite results so we use ACF PACF plots to see if the series is stationary
-acf(fresno_pm2_5_ts, main = "ACF of Original Time Series")
-pacf(fresno_pm2_5_ts, main = "PACF of Original Time Series")
+acf(pm2_5_ts, main = "ACF of Original Time Series")
+pacf(pm2_5_ts, main = "PACF of Original Time Series")
 
@@
 # Create weekly data
-fresno_pm2_5_df <- as.data.frame(fresno_pm2_5_ts)
-fresno_pm2_5_df$Date <- mdy(fresno_pm2_5$Date)
-colnames(fresno_pm2_5_df) <- c("PM2.5.AQI.Value", "Date")
+pm2_5_df <- as.data.frame(pm2_5_ts)
+pm2_5_df$Date <- mdy(pm2_5_data$Date)
+colnames(pm2_5_df) <- c("PM2.5.AQI.Value", "Date")
 
-fresno_pm2_5_xts <- as.xts(fresno_pm2_5_df, order.by=fresno_pm2_5_df$Date)
-fresno_pm2_5_xts <- fresno_pm2_5_xts[,-2]
-fresno_pm2_5_weekly <- apply.weekly(fresno_pm2_5_xts, mean)
+# Convert to xts format for weekly conversion
+pm2_5_xts <- as.xts(pm2_5_df, order.by=pm2_5_df$Date)
+pm2_5_xts <- pm2_5_xts[,-2]
+pm2_5_weekly <- apply.weekly(pm2_5_xts, mean)
+#pm2_5_weekly
 
-fresno_pm2_5_weekly_ts <- ts(fresno_pm2_5_weekly["19990103/20230805"], start = c(1999,1),frequency = 52.18)
-plot(fresno_pm2_5_weekly_ts)
+pm2_5_weekly_ts <- ts(pm2_5_weekly["19990103/20230811"], start = c(1999,1),frequency = 52.18)
+plot(pm2_5_weekly_ts)
 
-kpss.test(fresno_pm2_5_weekly_ts)
+kpss.test(pm2_5_weekly_ts)
 
-adf.test(fresno_pm2_5_weekly_ts)
+adf.test(pm2_5_weekly_ts)
 
-shapiro.test(fresno_pm2_5_weekly_ts)
-
-lambda <- BoxCox.lambda(fresno_pm2_5_weekly_ts)
-
-ts.plot(BoxCox(fresno_pm2_5_weekly_ts, lambda), main="Box-Cox Transformed PM 2.5 AQI")
-
-fresno_pm2_5_weekly_ts_transformed <- BoxCox(fresno_pm2_5_weekly_ts, lambda)
-
-kpss.test(fresno_pm2_5_weekly_ts_transformed)
-
-adf.test(fresno_pm2_5_weekly_ts_transformed)
-
-acf(fresno_pm2_5_weekly_ts_transformed, main = "ACF of Box-Cox Transformed Time Series")
-pacf(fresno_pm2_5_weekly_ts_transformed, main = "PACF of Box-Cox Transformed Time Series")
+acf(pm2_5_monthly_ts, main = "ACF of Weekly Time Series")
+pacf(pm2_5_monthly_ts, main = "PACF of Weekly Time Series")
 
@@
 # Create monthly data
-fresno_pm2_5_monthly <- apply.monthly(fresno_pm2_5_xts, mean)
+pm2_5_monthly <- apply.monthly(pm2_5_xts, mean)
+#pm2_5_monthly
 
-fresno_pm2_5_monthly_ts <- ts(fresno_pm2_5_monthly["19990130/20230805"], start = c(1999,1), frequency = 12)
-plot(fresno_pm2_5_monthly_ts)
+pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999,1), frequency = 12)
+plot(pm2_5_monthly_ts)
 
-kpss.test(fresno_pm2_5_monthly_ts)
+kpss.test(pm2_5_monthly_ts)
 
-adf.test(fresno_pm2_5_monthly_ts)
+adf.test(pm2_5_monthly_ts)
 
-acf(fresno_pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
-pacf(fresno_pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
+acf(pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
+pacf(pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
+
 ```
+
+#### Step 5: Understand time series variance
+
+```{r, tidy=TRUE}
+
+plot(pm2_5_weekly_ts)
+
+# Strong seasonality, Strong cyclic, light downward trend, variance is reducing
+
+# Let's see how seasonality looks over time and is the variance changing
+pm2_5_df$Month <- month(pm2_5_df$Date)
+pm2_5_df$Year <- year(pm2_5_df$Date)
+
+avg_aqi_by_month_year <- pm2_5_df %>%
+  group_by(pm2_5_df$Year, pm2_5_df$Month) %>%
+  summarise(
+    avg_value = mean(PM2.5.AQI.Value)
+  )
+colnames(avg_aqi_by_month_year) <- c("Year", "Month", "avg_value")
+
+reshape_avg_aqi_by_month_year <- 
+  sqldf(
+    "SELECT 
+      Month,
+      MAX(CASE WHEN Year = 1999 THEN avg_value END) AS Year_1999,
+      MAX(CASE WHEN Year = 2000 THEN avg_value END) AS Year_2000,
+      MAX(CASE WHEN Year = 2001 THEN avg_value END) AS Year_2001,
+      MAX(CASE WHEN Year = 2002 THEN avg_value END) AS Year_2002,
+      MAX(CASE WHEN Year = 2003 THEN avg_value END) AS Year_2003,
+      MAX(CASE WHEN Year = 2004 THEN avg_value END) AS Year_2004,
+      MAX(CASE WHEN Year = 2005 THEN avg_value END) AS Year_2005,
+      MAX(CASE WHEN Year = 2006 THEN avg_value END) AS Year_2006,
+      MAX(CASE WHEN Year = 2007 THEN avg_value END) AS Year_2007,
+      MAX(CASE WHEN Year = 2008 THEN avg_value END) AS Year_2008,
+      MAX(CASE WHEN Year = 2009 THEN avg_value END) AS Year_2009,
+      MAX(CASE WHEN Year = 2010 THEN avg_value END) AS Year_2010,
+      MAX(CASE WHEN Year = 2011 THEN avg_value END) AS Year_2011,
+      MAX(CASE WHEN Year = 2012 THEN avg_value END) AS Year_2012,
+      MAX(CASE WHEN Year = 2013 THEN avg_value END) AS Year_2013,
+      MAX(CASE WHEN Year = 2014 THEN avg_value END) AS Year_2014,
+      MAX(CASE WHEN Year = 2015 THEN avg_value END) AS Year_2015,
+      MAX(CASE WHEN Year = 2016 THEN avg_value END) AS Year_2016,
+      MAX(CASE WHEN Year = 2017 THEN avg_value END) AS Year_2017,
+      MAX(CASE WHEN Year = 2018 THEN avg_value END) AS Year_2018,
+      MAX(CASE WHEN Year = 2019 THEN avg_value END) AS Year_2019,
+      MAX(CASE WHEN Year = 2020 THEN avg_value END) AS Year_2020,
+      MAX(CASE WHEN Year = 2021 THEN avg_value END) AS Year_2021,
+      MAX(CASE WHEN Year = 2022 THEN avg_value END) AS Year_2022,
+      MAX(CASE WHEN Year = 2023 THEN avg_value END) AS Year_2023
+    FROM avg_aqi_by_month_year
+    GROUP BY Month
+    ORDER BY Month"
+  )
+
+colnames(reshape_avg_aqi_by_month_year) <- c(
+  "Month", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", 
+  "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", 
+  "2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023")
+
+boxplot(reshape_avg_aqi_by_month_year[2:26])
+
+# We can see that over the years there is downward trend and variance is decreasing
+# except 2020 when we see a high peak likely caused by wildfires
+# Link: https://en.wikipedia.org/wiki/Lake_Fire_(2020)
+
+```
+
+#### Step 6: Split data into train and test
+
+```{r, tidy=TRUE}
+
+# Split the data into a training and test dataset
+train <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
+test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,6))
+
+```
+
+#### Step 7: Decompose the weekly plot
+
+```{r, tidy=TRUE}
+
+# Looking at spectral analysis
+periodogram(train, log = "no", plot=TRUE, 
+            ylab = "Periodogram",
+            xlab = "Frequency",
+            lwd=2, xlim = c(0, 0.06))
+
+# There are two trends: 
+#  1) A typical yearly (52 weeks) seasonal trend
+#  2) A trend that is repeating every 9.6 years (500 weeks)
+#  3) A typical half yearly (26 weeks) seasonal trend
+
+# Overall, there is no mixed strong trend that normal seasonality model cannot
+# capture.
+
+# Decompose the series
+plot(decompose(train, type="multiplicative"))
+
+# Important Inferences
+# 1) The PM2.5 AQI has been decreasing overall though there are rise every now and then.
+#    However the trend is going down with time.
+# 2) Winter months have a strong seasonal effect with Nov and Dec being the peak months
+#    usually which likely could be due to cold temperatures causing pollutant to
+#    not escape the lower atmosphere.
+#    Link: https://www.accuweather.com/en/health-wellness/why-air-pollution-is-worse-in-winter/689434#:~:text=Cold%20air%20is%20denser%20and%20moves%20slower%20than%20warm%20air,rate%20than%20during%20the%20summer.
+# 3) We can see a seasonal cycle of 12 months where the mean value of each month starts 
+#    with a increasing trend in the beginning of the year and drops down towards 
+#    the end of the year. We can see a seasonal effect with a cycle of 12 months.
+
+# Understand the seasonality and remove it to see the trend
+train_diff <- diff(train, lag = 52)
+autoplot(train_diff, ylab="Train Seasonal Differencing")
+
+# Let's look if the series is stationary?
+kpss.test(train_diff)
+# The reported p-value is 0.1, which is > 0.05, and would suggest that we fail 
+# to reject the null hypothesis of level stationarity (conclusion: stationary)
+
+adf.test(train_diff)
+# The reported p-value of 0.01 (<0.05) so we do reject the null hypothesis that 
+# the time series is non-stationary (conclusion: stationary)
+
+acf(train_diff, main = "ACF of Seasonal Differencing Time Series")
+pacf(train_diff, main = "PACF of Seasonal Differencing Time Series")
+
+```
+
+#### Step 6: Benchmark using snaive model
+
+```{r, tidy=TRUE}
+
+# Forecast the seasonal naïve for each month of 1980
+forecast_snaive <- snaive(train, h=110)
+# Plot the forecasts for snaive model
+plot(forecast_snaive, main = "PM 2.5 AQI Forecast - SNaive",
+         xlab = "Week", ylab = "PM 2.5 AQI")
+lines(test)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_snaive <- mean(abs((test - forecast_snaive$mean)/test))
+mape_snaive
+
+# Mean Squared Error (MSE)
+mse_snaive <- mean((test - forecast_snaive$mean)^2)
+mse_snaive
+
+```

### 2023-08-13 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

@@
 
-acf(pm2_5_monthly_ts, main = "ACF of Weekly Time Series")
-pacf(pm2_5_monthly_ts, main = "PACF of Weekly Time Series")
+acf(pm2_5_weekly_ts, main = "ACF of Weekly Time Series")
+pacf(pm2_5_weekly_ts, main = "PACF of Weekly Time Series")
 
@@
 
-#### Step 6: Split data into train and test
+#### Step 6: Split data into train and test - Weekly
 
@@
 train <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
-test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,6))
+test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,7))
 
@@
 
-#### Step 7: Decompose the weekly plot
+#### Step 7: Decompose the time series plot - Weekly
 
@@
 
-# Overall, there is no mixed strong trend that normal seasonality model cannot
-# capture.
+# Overall, there is no mixed effect that normal seasonality model cannot capture.
 
@@
 
-#### Step 6: Benchmark using snaive model
+#### Step 8: Benchmark using snaive model - Weekly
 
@@
 
-# Forecast the seasonal naïve for each month of 1980
+# Forecast the seasonal naïve for (2021-01 to 2023-07)
 forecast_snaive <- snaive(train, h=110)
@@
 ```
+
+#### Step 9: Split data into train and test - Monthly
+
+```{r, tidy=TRUE}
+
+# Split the data into a training and test dataset
+train_monthly <- window(pm2_5_monthly_ts, start = c(1999,1), end=c(2020,12))
+test_monthly <- window(pm2_5_monthly_ts, start = c(2021,1), end=c(2023,7))
+
+```
+
+#### Step 10: Decompose the time series plot - Monthly
+
+```{r, tidy=TRUE}
+
+# Looking at spectral analysis
+periodogram(train_monthly, log = "no", plot=TRUE, 
+            ylab = "Periodogram",
+            xlab = "Frequency",
+            lwd=2, xlim = c(0, 0.2))
+
+# There are two trends: 
+#  1) A typical yearly (12 months) seasonal trend
+#  2) A trend that is repeating every 8-9 years (100 months)
+#  3) A typical half yearly (6 months) seasonal trend
+
+# Overall, there is no mixed effect that normal seasonality model cannot capture.
+
+# Decompose the series
+plot(decompose(train_monthly, type="multiplicative"))
+
+# Important Inferences
+# 1) The PM2.5 AQI has been decreasing overall though there are rise every now and then.
+#    However the trend is going down with time.
+# 2) Winter months have a strong seasonal effect with Nov and Dec being the peak months
+#    usually which likely could be due to cold temperatures causing pollutant to
+#    not escape the lower atmosphere.
+#    Link: https://www.accuweather.com/en/health-wellness/why-air-pollution-is-worse-in-winter/689434#:~:text=Cold%20air%20is%20denser%20and%20moves%20slower%20than%20warm%20air,rate%20than%20during%20the%20summer.
+# 3) We can see a seasonal cycle of 12 months where the mean value of each month starts 
+#    with a increasing trend in the beginning of the year and drops down towards 
+#    the end of the year. We can see a seasonal effect with a cycle of 12 months.
+
+# Understand the seasonality and remove it to see the trend
+train_monthly_diff <- diff(train_monthly, lag = 12)
+autoplot(train_monthly_diff, ylab="Train Seasonal Differencing (Monthly)")
+
+# Let's look if the series is stationary?
+kpss.test(train_monthly_diff)
+# The reported p-value is 0.1, which is > 0.05, and would suggest that we fail 
+# to reject the null hypothesis of level stationarity (conclusion: stationary)
+
+adf.test(train_monthly_diff)
+# The reported p-value of 0.01 (<0.05) so we do reject the null hypothesis that 
+# the time series is non-stationary (conclusion: stationary)
+
+acf(train_monthly_diff, main = "ACF of Seasonal Differencing Time Series")
+pacf(train_monthly_diff, main = "PACF of Seasonal Differencing Time Series")
+
+```
+
+#### Step 11: Benchmark using snaive model - Monthly
+
+```{r, tidy=TRUE}
+
+# Forecast the seasonal naïve for (2021-01 to 2023-07)
+forecast_snaive_monthly <- snaive(train_monthly, h=31)
+# Plot the forecasts for snaive model
+plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
+         xlab = "Week", ylab = "PM 2.5 AQI")
+lines(test_monthly)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_snaive <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
+mape_snaive
+
+# Mean Squared Error (MSE)
+mse_snaive <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
+mse_snaive
+
+```

### 2023-08-13 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

@@
 ```
-
-#### Step 3: Extract weekly values for time series and plot the series
+#### Step 3: Prepare data for Weekly & Monthly extraction and EDA
 
 ```{r, tidy=TRUE}
-# Create weekly data
+# Create data with Dates
 pm2_5_df <- as.data.frame(pm2_5_ts)
@@
 
-# Convert to xts format for weekly conversion
+# Convert to xts format for weekly & monthly
 pm2_5_xts <- as.xts(pm2_5_df, order.by=pm2_5_df$Date)
 pm2_5_xts <- pm2_5_xts[,-2]
-pm2_5_weekly <- apply.weekly(pm2_5_xts, mean)
-#pm2_5_weekly
-
-pm2_5_weekly_ts <- ts(pm2_5_weekly["19990103/20230811"], start = c(1999,1),frequency = 52.18)
-plot(pm2_5_weekly_ts)
-
-kpss.test(pm2_5_weekly_ts)
-
-adf.test(pm2_5_weekly_ts)
-
-acf(pm2_5_weekly_ts, main = "ACF of Weekly Time Series")
-pacf(pm2_5_weekly_ts, main = "PACF of Weekly Time Series")
-
 ```
 
-#### Step 4: Extract monthly values for time series and plot the series
+#### Step 4: Understand PM2.5 AQI variance across years
 
 ```{r, tidy=TRUE}
-# Create monthly data
-pm2_5_monthly <- apply.monthly(pm2_5_xts, mean)
-#pm2_5_monthly
-
-pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999,1), frequency = 12)
-plot(pm2_5_monthly_ts)
-
-kpss.test(pm2_5_monthly_ts)
-
-adf.test(pm2_5_monthly_ts)
-
-acf(pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
-pacf(pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
-
-```
-
-#### Step 5: Understand time series variance
-
-```{r, tidy=TRUE}
-
-plot(pm2_5_weekly_ts)
-
-# Strong seasonality, Strong cyclic, light downward trend, variance is reducing
-
 # Let's see how seasonality looks over time and is the variance changing
@@
 # Link: https://en.wikipedia.org/wiki/Lake_Fire_(2020)
+```
+
+#### Step 5a: Extract values for time series and plot the series - Weekly
+
+```{r, tidy=TRUE}
+pm2_5_weekly <- apply.weekly(pm2_5_xts, mean)
+pm2_5_weekly_ts <- ts(pm2_5_weekly["19990103/20230811"], 
+                      start = c(1999,1),
+                      frequency = 52.18)
+plot(pm2_5_weekly_ts)
+# Strong seasonality, Strong cyclic, light downward trend, variance is reducing
+
+kpss.test(pm2_5_weekly_ts)
+
+adf.test(pm2_5_weekly_ts)
+
+acf(pm2_5_weekly_ts, main = "ACF of Weekly Time Series")
+pacf(pm2_5_weekly_ts, main = "PACF of Weekly Time Series")
 
@@
 
-#### Step 6: Split data into train and test - Weekly
+#### Step 6a: Split data into train and test - Weekly
 
 ```{r, tidy=TRUE}
-
 # Split the data into a training and test dataset
-train <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
-test <- window(pm2_5_weekly_ts, start = c(2021,1), end=c(2023,7))
-
+train_weekly <- window(pm2_5_weekly_ts, start = c(1999,1), end=c(2020,52))
+test_weekly <- window(pm2_5_weekly_ts, start = c(2021,1))
 ```
 
-#### Step 7: Decompose the time series plot - Weekly
+#### Step 7a: Decompose the time series plot - Weekly
 
 ```{r, tidy=TRUE}
-
 # Looking at spectral analysis
-periodogram(train, log = "no", plot=TRUE, 
+periodogram(train_weekly, log = "no", plot=TRUE, 
             ylab = "Periodogram",
@@
 # Decompose the series
-plot(decompose(train, type="multiplicative"))
+plot(decompose(train_weekly, type="multiplicative"))
 
@@
 # Understand the seasonality and remove it to see the trend
-train_diff <- diff(train, lag = 52)
-autoplot(train_diff, ylab="Train Seasonal Differencing")
+train_weekly_diff <- diff(train_weekly, lag = 52)
+autoplot(train_weekly_diff, ylab="Train Seasonal Differencing - Weekly Data")
 
 # Let's look if the series is stationary?
-kpss.test(train_diff)
+kpss.test(train_weekly_diff)
 # The reported p-value is 0.1, which is > 0.05, and would suggest that we fail 
@@
 
-adf.test(train_diff)
+adf.test(train_weekly_diff)
 # The reported p-value of 0.01 (<0.05) so we do reject the null hypothesis that 
@@
 
-acf(train_diff, main = "ACF of Seasonal Differencing Time Series")
-pacf(train_diff, main = "PACF of Seasonal Differencing Time Series")
+acf(train_weekly_diff, main = "ACF of Seasonal Differencing Time Series - Weekly Data")
+pacf(train_weekly_diff, main = "PACF of Seasonal Differencing Time Series - Weekly Data")
 
@@
 
-#### Step 8: Benchmark using snaive model - Weekly
+#### Step 8a: Benchmark using snaive model - Weekly
 
 ```{r, tidy=TRUE}
-
 # Forecast the seasonal naïve for (2021-01 to 2023-07)
-forecast_snaive <- snaive(train, h=110)
+forecast_snaive_weekly <- snaive(train_weekly, h=110)
 # Plot the forecasts for snaive model
-plot(forecast_snaive, main = "PM 2.5 AQI Forecast - SNaive",
+plot(forecast_snaive_weekly, main = "PM 2.5 AQI Forecast - SNaive (Weekly)",
          xlab = "Week", ylab = "PM 2.5 AQI")
-lines(test)
+lines(test_weekly)
 
@@
 # Mean Absolute Percentage Error (MAPE)
-mape_snaive <- mean(abs((test - forecast_snaive$mean)/test))
-mape_snaive
+mape_snaive_weekly <- mean(abs((test_weekly - forecast_snaive_weekly$mean)/test_weekly))
+mape_snaive_weekly
+
+# Mean Absolute Error (MAE)
+mae_snaive_weekly <- mean(abs((test_weekly - forecast_snaive_weekly$mean)))
+mae_snaive_weekly
 
 # Mean Squared Error (MSE)
-mse_snaive <- mean((test - forecast_snaive$mean)^2)
-mse_snaive
+mse_snaive_weekly <- mean((test_weekly - forecast_snaive_weekly$mean)^2)
+mse_snaive_weekly
+
+# Evaluate the residuals
+checkresiduals(forecast_snaive_weekly)
+
+```
+#### Step 5b: Extract values for time series and plot the series - Monthly
+
+```{r, tidy=TRUE}
+# Create monthly data
+pm2_5_monthly <- apply.monthly(pm2_5_xts, mean)
+
+pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999,1), frequency = 12)
+plot(pm2_5_monthly_ts)
+
+kpss.test(pm2_5_monthly_ts)
+
+adf.test(pm2_5_monthly_ts)
+
+acf(pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
+pacf(pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
 
@@
 
-#### Step 9: Split data into train and test - Monthly
+#### Step 6b: Split data into train and test - Monthly
 
 ```{r, tidy=TRUE}
-
 # Split the data into a training and test dataset
 train_monthly <- window(pm2_5_monthly_ts, start = c(1999,1), end=c(2020,12))
-test_monthly <- window(pm2_5_monthly_ts, start = c(2021,1), end=c(2023,7))
-
+test_monthly <- window(pm2_5_monthly_ts, start = c(2021,1))
 ```
 
-#### Step 10: Decompose the time series plot - Monthly
+#### Step 7b: Decompose the time series plot - Monthly
 
@@
 
-#### Step 11: Benchmark using snaive model - Monthly
+#### Step 8b: Benchmark using snaive model - Monthly
 
@@
 
+# Evaluate the residuals
+checkresiduals(forecast_snaive_monthly)
+
 ```

### 2023-08-13 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
Add files via upload

--- 2023-08-13 MSCA_31006_Time_Series_Final_Project.Rmd
@@
 ```{r, tidy=TRUE}
-path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
-pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
+#path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
+#pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
+#                     na.strings=c("", "NA"))
+#ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
+#                     na.strings=c("", "NA"))
+
+path <- "D:/TS Final/"
+pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
                      na.strings=c("", "NA"))
-ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
+ozone_data <- read.csv(paste0(path, "ozone_LA.csv"), 
                      na.strings=c("", "NA"))
@@
 ```
+
+#### Step 9a: TBATS model - Weekly
+
+```{r, tidy=TRUE}
+vec <- as.vector(train_weekly)
+
+univariate_ts <- ts(vec, start=1999, frequency=52.18)
+
+stl_decomposition <- stl(univariate_ts, s.window="periodic")
+plot(stl_decomposition)
+
+h <- length(test_weekly)
+
+model_tbats <- tbats(train_weekly)
+forecast_tbats <- forecast(model_tbats, h=h)
+plot(forecast_tbats)
+
+checkresiduals(model_tbats)
+#Ljung-Box test .05179
+
+plot(pm2_5_weekly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_weekly_ts), max(pm2_5_weekly_ts)), main="Train, Test, and Forecasted Data - Weekly", xlab="Year", ylab="PM2.5 AQI Value")
+lines(test_weekly, col="red") # Test data in red
+lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
+
+mae <- mean(abs(test_weekly - forecast_tbats$mean))
+print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
+#Mean Absolute Error (MAE): 9.31
+
+comparison_df <- data.frame(
+  Date = time(test_weekly),
+  Actual = as.vector(test_weekly),
+  Forecasted = as.vector(forecast_tbats$mean)
+)
+print(comparison_df)
+
+```
+
+#### Step 9b: TBATS model - Monthly
+
+```{r, tidy=TRUE}
+
+vec <- as.vector(train_monthly)
+
+univariate_ts <- ts(vec, start=1999, frequency=52.18)
+
+stl_decomposition <- stl(univariate_ts, s.window="periodic")
+plot(stl_decomposition)
+
+h <- length(test_monthly)
+
+model_tbats <- tbats(train_monthly)
+forecast_tbats <- forecast(model_tbats, h=h)
+plot(forecast_tbats)
+
+checkresiduals(model_tbats)
+#Ljung-Box test .07956
+
+plot(pm2_5_monthly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_ts), max(pm2_5_monthly_ts)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
+lines(test_monthly, col="red") # Test data in red
+lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
+
+mae <- mean(abs(test_monthly - forecast_tbats$mean))
+print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
+#Mean Absolute Error (MAE): 6.61
+
+comparison_df <- data.frame(
+  Date = time(test_monthly),
+  Actual = as.vector(test_monthly),
+  Forecasted = as.vector(forecast_tbats$mean)
+)
+print(comparison_df)
+
+```

### 2023-08-14 MSCA_31006_Time_Series_Final_Project - ARIMA.Rmd
ARIMA

--- 2023-08-13 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
@@
 
-path <- "D:/TS Final/"
+path <- "/Users/sydneypeters/Desktop/"
 pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
@@
 ```
+
+#### ARIMA model using auto.arima() - Weekly
+```{r}
+# Convert weekly data to a time series object
+pm2_5_weekly_ts <- ts(pm2_5_weekly["19990103/20230811"], start = c(1999, 1), frequency = 52.18)
+
+# Check stationarity, perform differencing if needed
+kpss_test_result <- kpss.test(pm2_5_weekly_ts)
+adf_test_result <- adf.test(pm2_5_weekly_ts)
+
+if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
+  pm2_5_weekly_diff <- diff(pm2_5_weekly_ts)
+} else {
+  pm2_5_weekly_diff <- pm2_5_weekly_ts
+}
+
+# Fit ARIMA model using auto.arima()
+arima_model_weekly <- auto.arima(pm2_5_weekly_diff)
+
+# Forecast using ARIMA model
+forecast_arima_weekly <- forecast(arima_model_weekly, h = h)
+plot(forecast_arima_weekly)
+```
+#### ARIMA model using auto.arima() - Monthly
+```{r}
+# Convert monthly data to a time series object
+pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999, 1), frequency = 12)
+
+# Check stationarity, perform differencing if needed
+kpss_test_result <- kpss.test(pm2_5_monthly_ts)
+adf_test_result <- adf.test(pm2_5_monthly_ts)
+
+if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
+  pm2_5_monthly_diff <- diff(pm2_5_monthly_ts)
+} else {
+  pm2_5_monthly_diff <- pm2_5_monthly_ts
+}
+
+# Fit ARIMA model using auto.arima()
+arima_model_monthly <- auto.arima(pm2_5_monthly_diff)
+
+# Forecast using ARIMA model
+forecast_arima_monthly <- forecast(arima_model_monthly, h = h)
+plot(forecast_arima_monthly)
+```

### 2023-08-14 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

@@
 ```
+
+#### Step 10a: Forecast using Neural Network - Monthly
+
+```{r, tidy=TRUE}
+
+model_nn_weekly <- nnetar(train_weekly,
+                           size=5, 
+                           decay=0.5,
+                           repeats=1000, 
+                           lambda="auto")
+summary(model_nn_weekly)
+forecast_nn_weekly <- forecast(model_nn_weekly, PI=FALSE, h=110)
+
+# Plot the forecasts for NN model
+plot(forecast_nn_weekly, main = "PM 2.5 AQI Forecast - Neural Network",
+         xlab = "Weekly", ylab = "PM 2.5 AQI")
+lines(test_weekly)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_nn_weekly <- mean(abs((test_weekly - forecast_nn_weekly$mean)/test_weekly))
+mape_nn_weekly
+
+# Mean Absolute Error (MAE)
+mae_nn_weekly <- mean(abs((test_weekly - forecast_nn_weekly$mean)))
+mae_nn_weekly
+
+# Mean Squared Error (MSE)
+mse_nn_weekly <- mean((test_weekly - forecast_nn_weekly$mean)^2)
+mse_nn_weekly
+
+# Evaluate the residuals
+checkresiduals(forecast_nn_weekly)
+```
+
 #### Step 5b: Extract values for time series and plot the series - Monthly
@@
 plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
-         xlab = "Week", ylab = "PM 2.5 AQI")
+         xlab = "Month", ylab = "PM 2.5 AQI")
 lines(test_monthly)
@@
 # Mean Absolute Percentage Error (MAPE)
-mape_snaive <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
-mape_snaive
+mape_snaive_monthly <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
+mape_snaive_monthly
+
+# Mean Absolute Error (MAE)
+mae_snaive_monthly <- mean(abs((test_monthly - forecast_snaive_monthly$mean)))
+mae_snaive_monthly
 
 # Mean Squared Error (MSE)
-mse_snaive <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
-mse_snaive
+mse_snaive_monthly <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
+mse_snaive_monthly
 
@@
 ```
+
+#### Step 9b: Forecast using ETS - Monthly
+
+```{r, tidy=TRUE}
+# Build the exponential smoothing model
+model_ets_monthly <- ets(train_monthly, lambda="auto")
+summary(model_ets_monthly)
+forecast_ets_monthly <- forecast(model_ets_monthly, h=31)
+
+# Plot the forecasts for ETS model
+plot(forecast_ets_monthly, main = "PM 2.5 AQI Forecast - ETS",
+         xlab = "Month", ylab = "PM 2.5 AQI")
+lines(test_monthly)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_ets_monthly <- mean(abs((test_monthly - forecast_ets_monthly$mean)/test_monthly))
+mape_ets_monthly
+
+# Mean Absolute Error (MAE)
+mae_ets_monthly <- mean(abs((test_monthly - forecast_ets_monthly$mean)))
+mae_ets_monthly
+
+# Mean Squared Error (MSE)
+mse_ets_monthly <- mean((test_monthly - forecast_ets_monthly$mean)^2)
+mse_ets_monthly
+
+# Evaluate the residuals
+checkresiduals(forecast_ets_monthly)
+
+Box.test(residuals(forecast_ets_monthly), lag = 12, type = "Ljung-Box")
+
+```
+
+#### Step 10b: Forecast using Neural Network - Monthly
+
+```{r, tidy=TRUE}
+
+model_nn_monthly <- nnetar(train_monthly,
+                           P=12,
+                           size=5, 
+                           decay=0.5,
+                           repeats=1000, 
+                           lambda="auto")
+summary(model_nn_monthly)
+forecast_nn_monthly <- forecast(model_nn_monthly, PI=FALSE, h=31)
+
+# Plot the forecasts for NN model
+plot(forecast_nn_monthly, main = "PM 2.5 AQI Forecast - Neural Network",
+         xlab = "Month", ylab = "PM 2.5 AQI")
+lines(test_monthly)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_nn_monthly <- mean(abs((test_monthly - forecast_nn_monthly$mean)/test_monthly))
+mape_nn_monthly
+
+# Mean Absolute Error (MAE)
+mae_nn_monthly <- mean(abs((test_monthly - forecast_nn_monthly$mean)))
+mae_nn_monthly
+
+# Mean Squared Error (MSE)
+mse_nn_monthly <- mean((test_monthly - forecast_nn_monthly$mean)^2)
+mse_nn_monthly
+
+# Evaluate the residuals
+checkresiduals(forecast_nn_monthly)
+```

### 2023-08-14 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

@@
 
+#### Step 9a: Forecast using ETS - Weekly
+
+```{r, tidy=TRUE}
+# Build the exponential smoothing model
+model_ets_weekly <- ets(train_weekly, lambda="auto", additive.only=TRUE)
+summary(model_ets_weekly)
+forecast_ets_weekly <- forecast(model_ets_weekly, h=110)
+
+# Plot the forecasts for ETS model
+plot(forecast_ets_weekly, main = "PM 2.5 AQI Forecast - ETS",
+         xlab = "Month", ylab = "PM 2.5 AQI")
+lines(test_weekly)
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Mean Absolute Percentage Error (MAPE)
+mape_ets_weekly <- mean(abs((test_weekly - forecast_ets_weekly$mean)/test_weekly))
+mape_ets_weekly
+
+# Mean Absolute Error (MAE)
+mae_ets_weekly <- mean(abs((test_weekly - forecast_ets_weekly$mean)))
+mae_ets_weekly
+
+# Mean Squared Error (MSE)
+mse_ets_weekly <- mean((test_weekly - forecast_ets_weekly$mean)^2)
+mse_ets_weekly
+
+# Evaluate the residuals
+checkresiduals(forecast_ets_weekly)
+
+Box.test(residuals(forecast_ets_weekly), lag = 52, type = "Ljung-Box")
+
+```
+
 #### Step 10a: Forecast using Neural Network - Monthly

### 2023-08-15 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
Add files via upload

@@
 # Compare the forecast with the actual test data by calculating MAPE and MSE
+mae_snaive_monthly <- mean(abs((test_monthly - forecast_snaive_monthly$mean)))
+mae_snaive_monthly
+
 # Mean Absolute Percentage Error (MAPE)
@@
 #Mean Absolute Error (MAE): 9.31 
+mape <- mean(abs(test - forecast_tbats$mean)/test * 100)
+cat("Mean Absolute Percentage Error (MAPE):", round(mape, 2), "\n")
 
@@
 print(comparison_df)
+
+model_tbats$lambda
+model_tbats$alpha
+model_tbats$beta
+model_tbats$seasonal.periods
+model_tbats$p
+model_tbats$q
+model_tbats$likelihood
+model_tbats$AIC
 

### 2023-08-15 MSCA_31006_Time_Series_Final_Project - ARIMA.Rmd
Add files via upload

--- 2023-08-13 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
@@
 ---
+```{r}
+# List all objects in the environment
+objects_to_remove <- ls()
+
+# Remove all objects
+rm(list = objects_to_remove)
+```
 
@@
 ```{r, tidy=TRUE}
-#path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
-#pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
+path <- "/Users/nikithagopal/Desktop/MSCA Data/Time Series Final/"
+pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
+                     na.strings=c("", "NA"))
+ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
+                     na.strings=c("", "NA"))
+
+#path <- "/Users/sydneypeters/Desktop/"
+#pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
 #                     na.strings=c("", "NA"))
-#ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
+#ozone_data <- read.csv(paste0(path, "ozone_LA.csv"), 
 #                     na.strings=c("", "NA"))
-
-path <- "D:/TS Final/"
-pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
-                     na.strings=c("", "NA"))
-ozone_data <- read.csv(paste0(path, "ozone_LA.csv"), 
-                     na.strings=c("", "NA"))
 
@@
 ```
+
 #### Step 5b: Extract values for time series and plot the series - Monthly
@@
 ```
+
+#### Step 10a: ARIMA model using auto.arima() - Weekly
+```{r}
+eacf <-
+function (z,ar.max=7,ma.max=13) 
+{
+#
+#  PROGRAMMED BY K.S. CHAN, DEPARTMENT OF STATISTICS AND ACTUARIAL SCIENCE,
+#  UNIVERSITY OF IOWA.
+#
+#  DATE: 4/2001
+#  Compute the extended sample acf (ESACF) for the time series stored in z.
+#  The matrix of ESACF with the AR order up to ar.max and the MA order
+#  up to ma.max is stored in the matrix EACFM.
+#  The default values for NAR and NMA are 7 and 13 respectively.
+#  Side effect of the eacf function:
+#  The function prints a coded ESACF table with
+#  significant values denoted by * and nosignificant values by 0, significance
+#  level being 5%.
+#
+#  Output:
+#   eacf=matrix of esacf
+#   symbol=matrix of coded esacf
+#
+
+lag1<-function(z,lag=1){c(rep(NA,lag),z[1:(length(z)-lag)])}
+reupm<-function(m1,nrow,ncol){
+k<-ncol-1
+m2<-NULL
+for (i in 1:k){
+i1<-i+1
+work<-lag1(m1[,i])
+work[1]<--1
+temp<-m1[,i1]-work*m1[i1,i1]/m1[i,i]
+temp[i1]<-0
+m2<-cbind(m2,temp)
+}
+m2}
+ceascf<-function(m,cov1,nar,ncol,count,ncov,z,zm){
+result<-0*seq(1,nar+1)
+result[1]<-cov1[ncov+count]
+for (i in 1:nar) {
+temp<-cbind(z[-(1:i)],zm[-(1:i),1:i])%*%c(1,-m[1:i,i])
+result[i+1]<-acf(temp,plot=FALSE,lag.max=count,drop.lag.0=FALSE)$acf[count+1]
+}
+result
+}
+
+ar.max<-ar.max+1
+ma.max<-ma.max+1
+nar<-ar.max-1
+nma<-ma.max
+ncov<-nar+nma+2
+nrow<-nar+nma+1
+ncol<-nrow-1
+z<-z-mean(z)
+zm<-NULL
+for(i in 1:nar) zm<-cbind(zm,lag1(z,lag=i))
+cov1<-acf(z,lag.max=ncov,plot=FALSE,drop.lag.0=FALSE)$acf
+cov1<-c(rev(cov1[-1]),cov1)
+ncov<-ncov+1
+m1<-matrix(0,ncol=ncol,nrow=nrow)
+for(i in 1:ncol) m1[1:i,i]<-
+ar.ols(z,order.max=i,aic=FALSE,demean=FALSE,intercept=FALSE)$ar
+eacfm<-NULL
+for (i in 1:nma) {
+m2<-reupm(m1=m1,nrow=nrow,ncol=ncol)
+ncol<-ncol-1
+eacfm<-cbind(eacfm, ceascf(m2,cov1,nar,ncol,i,ncov,z,zm))
+m1<-m2}
+work<-1:(nar+1)
+work<-length(z)-work+1
+symbol<-NULL
+for ( i in 1:nma) {
+work<-work-1
+symbol<-cbind(symbol,ifelse(abs(eacfm[,i])>2/work^.5, 'x','o'))}
+rownames(symbol)<-0:(ar.max-1)
+colnames(symbol)<-0:(ma.max-1)
+cat('AR/MA\n')
+print(symbol,quote=FALSE)
+invisible(list(eacf=eacfm,ar.max=ar.max,ma.ma=ma.max,symbol=symbol))
+}
+```
+
+```{r}
+eacf(train_weekly)
+# p <- 1,2,3
+# q <- 2,3,4
+```
+
+```{r}
+
+kpss_test_result <- kpss.test(train_weekly_diff)
+adf_test_result <- adf.test(train_weekly_diff)
+
+cat("KPSS: ", round(kpss_test_result$p.value, 2), "\n")
+cat("ADF: ", round(adf_test_result$p.value, 2), "\n")
+
+# Fit ARIMA model using auto.arima()
+# Find appropriate values for d and D using ndiffs and nsdiffs
+# d <- ndiffs(train_weekly_diff)
+# D <- nsdiffs(train_weekly_diff)
+
+# Fit ARIMA model using auto.arima() with seasonal components and previously found d and D
+arima_model_weekly <- auto.arima(train_weekly, seasonal = TRUE, stepwise = FALSE,
+                                 approximation = FALSE, trace = FALSE, lambda = "auto",
+                                 start.p = 1, max.p = 4,
+                                 start.q = 2, max.q = 5, 
+                                 d=0)
+
+# auto.arima(pm2_5_weekly_diff, lambda = "auto", seasonal=TRUE, )
+
+# Forecast using ARIMA model
+h <- length(test_weekly)
+forecast_arima_weekly <- forecast(arima_model_weekly, h = h)
+
+plot(forecast_arima_weekly, main = "PM 2.5 AQI Forecast - sARIMA",
+         xlab = "Week", ylab = "PM 2.5 AQI")
+lines(test_weekly)
+
+```
+```{r}
+forecast_arima_weekly
+```
+
+```{r}
+
+# Check the residuals of the ARIMA model
+checkresiduals(arima_model_weekly)
+
+# Calculate MAE
+mae <- mean(abs(forecast_arima_weekly$mean - test_weekly))
+# Calculate MAPE
+mape <- mean(abs((forecast_arima_weekly$mean - test_weekly) / test_weekly)) * 100
+
+# Get AICc and BIC values from the ARIMA model
+aic <- arima_model_weekly$aic
+bic <- arima_model_weekly$bic
+
+# Corrected AICc calculation
+n <- length(train_weekly)
+k <- length(arima_model_weekly$coef)
+aic_corrected <- aic + 2 * (k + 1) * (k + 2) / (n - k - 2)
+
+# Print the results
+cat("MAE: ", round(mae, 2), "\n")
+cat("MAPE: ", round(mape, 2), "%\n")
+cat("AICc: ", round(aic_corrected, 2), "\n")
+cat("BIC: ", round(bic, 2), "\n")
+
+# Create a time index for the forecasted values
+forecast_time_index <- seq(from = end(test_weekly) + 1 / 52, 
+                           length.out = length(forecasted_values), by = 1 / 52)
+
+# Plot historical data, actual test, and forecast
+plot(train_weekly, xlim=c(1999, 2023), ylim=c(min(train_weekly), max(train_weekly)), main="Train, Test, and Forecasted Data - Weekly", xlab="Year", ylab="PM2.5 AQI Value")
+lines(test_weekly, col="red") # Test data in red
+lines(forecast_arima_weekly$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
+
+comparison_df <- data.frame(
+  Date = time(test_weekly),
+  Actual = as.vector(test_weekly),
+  Forecasted = as.vector(forecast_arima_weekly$mean)
+)
+print(comparison_df)
+```
+```{r}
+summary(arima_model_weekly)
+```
+
+```{r}
+# Assuming your residuals are stored in a variable called "residuals"
+shapiro_test_result <- shapiro.test(arima_model_weekly$residuals)
+
+# Print the test result
+print(shapiro_test_result)
+
+# Check the p-value
+if (shapiro_test_result$p.value < 0.05) {
+  cat("The residuals are not normally distributed (p-value:", shapiro_test_result$p.value, ")\n")
+} else {
+  cat("The residuals appear to be normally distributed (p-value:", shapiro_test_result$p.value, ")\n")
+}
+```
+
+#### ARIMA model using auto.arima() - Monthly
+
+```{r}
+# Convert monthly data to a time series object
+# pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999, 1), frequency = 12)
+
+y_values_monthly <- train_monthly
+optimal_lambda <- BoxCox.lambda(y_values_monthly)
+
+# Apply the Box-Cox transformation
+transformed_data_monthly <- if (optimal_lambda != 0) {
+  (y_values_monthly^optimal_lambda - 1) / optimal_lambda
+} else {
+  log(y_values_monthly)
+}
+
+# Check stationarity, perform differencing if needed
+kpss_test_result <- kpss.test(pm2_5_monthly_ts)
+adf_test_result <- adf.test(pm2_5_monthly_ts)
+
+if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
+  pm2_5_monthly_diff <- diff(pm2_5_monthly_ts)
+} else {
+  pm2_5_monthly_diff <- pm2_5_monthly_ts
+}
+
+# Check stationarity, perform differencing if needed
+kpss_test_result <- kpss.test(transformed_data_monthly)
+adf_test_result <- adf.test(transformed_data_monthly)
+
+if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
+  pm2_5_monthly_diff <- diff(transformed_data_monthly)
+
+  # Check stationarity after first differencing
+  kpss_test_result_diff <- kpss.test(pm2_5_monthly_diff)
+  adf_test_result_diff <- adf.test(pm2_5_monthly_diff)
+
+  if (kpss_test_result_diff$p.value < 0.05 || adf_test_result_diff$p.value < 0.05) {
+    # If still non-stationary, apply second-order differencing
+    pm2_5_monthly_diff <- diff(pm2_5_monthly_diff)
+
+    print("First and Second Order Differencing was performed")
+  }
+} else {
+  pm2_5_monthly_diff <- transformed_data_monthly
+  print("First Order Differencing was performed")
+}
+
+# Fit ARIMA model using auto.arima()
+arima_model_monthly <- auto.arima(pm2_5_monthly_diff, seasonal = TRUE)
+
+# Forecast using ARIMA model
+h <- length(test_monthly)
+forecast_arima_monthly <- forecast(arima_model_monthly, h = h)
+
+original_scale_forecasts_monthly <- InvBoxCox(forecast_arima_monthly$mean, lambda = optimal_lambda)
+
+print(cat("Optimal Lambda Value:", optimal_lambda, "\n"))
+
+# Plot the original data
+original_scale_forecasts_monthly <- original_scale_forecasts_monthly * 100
+plot(train_monthly, main = "Original Data and Forecasts")
+lines(original_scale_forecasts_monthly, col = "blue")  # Add forecast line in blue
+
+# Add legend
+legend("topleft", legend = c("Original Data", "Forecasts"), col = c("black", "blue"), lwd = 1)
+
+```
+```{r}
+
+checkresiduals(arima_model_monthly)
+#Ljung-Box test .05179
+
+plot(pm2_5_monthly_diff, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_diff), max(pm2_5_monthly_diff)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
+lines(test_monthly, col="red") # Test data in red
+lines(original_scale_forecasts_monthly, col="blue", type="o") # Forecasted data in blue
+legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
+
+mae <- mean(abs(test_monthly - original_scale_forecasts_monthly))
+print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
+#Mean Absolute Error (MAE): 9.31
+
+comparison_df <- data.frame(
+  Date = time(test_monthly),
+  Actual = as.vector(test_monthly),
+  Forecasted = as.vector(original_scale_forecasts_monthly)
+)
+print(comparison_df)
+
+```

### 2023-08-16 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
Add files via upload

@@
 #Mean Absolute Error (MAE): 9.31 
-mape <- mean(abs(test - forecast_tbats$mean)/test * 100)
-cat("Mean Absolute Percentage Error (MAPE):", round(mape, 2), "\n")
+smape_tbats_weekly <- smape(test_weekly, forecast_tbats$mean)
+print(cat("Symmetric Mean Absolute Percentage Error (MAPE):", round(smape_tbats_weekly, 4), "\n"))
+rmse <- rmse(test_weekly, forecast_tbats$mean)
+print(cat("Root Mean Squared Error (RMSE):", round(rmse, 2), "\n"))
 
@@
 model_tbats$seasonal.periods
+model_tbats$ar.coefficients
+model_tbats$ma.coefficients
 model_tbats$p

### 2023-08-16 model/SARIMA.Rmd
Rename MSCA_31006_Time_Series_Final_Project - ARIMA.Rmd to SARIMA.Rmd

--- 2023-08-15 MSCA_31006_Time_Series_Final_Project - ARIMA.Rmd
@@
 # Fit ARIMA model using auto.arima()
-# Find appropriate values for d and D using ndiffs and nsdiffs
-# d <- ndiffs(train_weekly_diff)
-# D <- nsdiffs(train_weekly_diff)
-
-# Fit ARIMA model using auto.arima() with seasonal components and previously found d and D
 arima_model_weekly <- auto.arima(train_weekly, seasonal = TRUE, stepwise = FALSE,
-                                 approximation = FALSE, trace = FALSE, lambda = "auto",
+                                 trace = FALSE,
                                  start.p = 1, max.p = 4,
-                                 start.q = 2, max.q = 5, 
-                                 d=0)
+                                 start.q = 2, max.q = 5)
 
@@
 
-```
-```{r}
-forecast_arima_weekly
 ```
@@
 ```
-
-#### ARIMA model using auto.arima() - Monthly
-
-```{r}
-# Convert monthly data to a time series object
-# pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999, 1), frequency = 12)
-
-y_values_monthly <- train_monthly
-optimal_lambda <- BoxCox.lambda(y_values_monthly)
-
-# Apply the Box-Cox transformation
-transformed_data_monthly <- if (optimal_lambda != 0) {
-  (y_values_monthly^optimal_lambda - 1) / optimal_lambda
-} else {
-  log(y_values_monthly)
-}
-
-# Check stationarity, perform differencing if needed
-kpss_test_result <- kpss.test(pm2_5_monthly_ts)
-adf_test_result <- adf.test(pm2_5_monthly_ts)
-
-if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
-  pm2_5_monthly_diff <- diff(pm2_5_monthly_ts)
-} else {
-  pm2_5_monthly_diff <- pm2_5_monthly_ts
-}
-
-# Check stationarity, perform differencing if needed
-kpss_test_result <- kpss.test(transformed_data_monthly)
-adf_test_result <- adf.test(transformed_data_monthly)
-
-if (kpss_test_result$p.value < 0.05 || adf_test_result$p.value < 0.05) {
-  pm2_5_monthly_diff <- diff(transformed_data_monthly)
-
-  # Check stationarity after first differencing
-  kpss_test_result_diff <- kpss.test(pm2_5_monthly_diff)
-  adf_test_result_diff <- adf.test(pm2_5_monthly_diff)
-
-  if (kpss_test_result_diff$p.value < 0.05 || adf_test_result_diff$p.value < 0.05) {
-    # If still non-stationary, apply second-order differencing
-    pm2_5_monthly_diff <- diff(pm2_5_monthly_diff)
-
-    print("First and Second Order Differencing was performed")
-  }
-} else {
-  pm2_5_monthly_diff <- transformed_data_monthly
-  print("First Order Differencing was performed")
-}
-
-# Fit ARIMA model using auto.arima()
-arima_model_monthly <- auto.arima(pm2_5_monthly_diff, seasonal = TRUE)
-
-# Forecast using ARIMA model
-h <- length(test_monthly)
-forecast_arima_monthly <- forecast(arima_model_monthly, h = h)
-
-original_scale_forecasts_monthly <- InvBoxCox(forecast_arima_monthly$mean, lambda = optimal_lambda)
-
-print(cat("Optimal Lambda Value:", optimal_lambda, "\n"))
-
-# Plot the original data
-original_scale_forecasts_monthly <- original_scale_forecasts_monthly * 100
-plot(train_monthly, main = "Original Data and Forecasts")
-lines(original_scale_forecasts_monthly, col = "blue")  # Add forecast line in blue
-
-# Add legend
-legend("topleft", legend = c("Original Data", "Forecasts"), col = c("black", "blue"), lwd = 1)
-
-```
-```{r}
-
-checkresiduals(arima_model_monthly)
-#Ljung-Box test .05179
-
-plot(pm2_5_monthly_diff, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_diff), max(pm2_5_monthly_diff)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
-lines(test_monthly, col="red") # Test data in red
-lines(original_scale_forecasts_monthly, col="blue", type="o") # Forecasted data in blue
-legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
-
-mae <- mean(abs(test_monthly - original_scale_forecasts_monthly))
-print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
-#Mean Absolute Error (MAE): 9.31
-
-comparison_df <- data.frame(
-  Date = time(test_monthly),
-  Actual = as.vector(test_monthly),
-  Forecasted = as.vector(original_scale_forecasts_monthly)
-)
-print(comparison_df)
-
-```

### 2023-08-16 MSCA_31006_Time_Series_Final_Combined.Rmd
Add files via upload

--- 2023-08-16 model/SARIMA.Rmd
@@
 ---
-```{r}
-# List all objects in the environment
-objects_to_remove <- ls()
-
-# Remove all objects
-rm(list = objects_to_remove)
-```
 
@@
 ```{r, tidy=TRUE}
-path <- "/Users/nikithagopal/Desktop/MSCA Data/Time Series Final/"
+path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
 pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
@@
                      na.strings=c("", "NA"))
-
-#path <- "/Users/sydneypeters/Desktop/"
-#pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
-#                     na.strings=c("", "NA"))
-#ozone_data <- read.csv(paste0(path, "ozone_LA.csv"), 
-#                     na.strings=c("", "NA"))
 
@@
 
-#### Step 6a: Split data into train and test - Weekly
+#### Step 6: Split data into train and test - Weekly
 
@@
 
-#### Step 7a: Decompose the time series plot - Weekly
+#### Step 7: Decompose the time series plot - Weekly
 
@@
 
-#### Step 8a: Benchmark using snaive model - Weekly
+#### Step 8: Benchmark using snaive model - Weekly
 
@@
 # Forecast the seasonal naïve for (2021-01 to 2023-07)
-forecast_snaive_weekly <- snaive(train_weekly, h=110)
+h <- length(test_weekly)
+forecast_snaive_weekly <- snaive(train_weekly, lambda="auto", h)
+summary(forecast_snaive_weekly)
 # Plot the forecasts for snaive model
@@
 # Compare the forecast with the actual test data by calculating MAPE and MSE
-# Mean Absolute Percentage Error (MAPE)
-mape_snaive_weekly <- mean(abs((test_weekly - forecast_snaive_weekly$mean)/test_weekly))
-mape_snaive_weekly
+# Symmetric Mean Absolute Percentage Error (sMAPE)
+smape_snaive_weekly <- smape(test_weekly, forecast_snaive_weekly$mean)
+smape_snaive_weekly
 
@@
 
-# Mean Squared Error (MSE)
-mse_snaive_weekly <- mean((test_weekly - forecast_snaive_weekly$mean)^2)
-mse_snaive_weekly
+# Root Mean Squared Error (RMSE)
+rmse_snaive_weekly <- rmse(test_weekly, forecast_snaive_weekly$mean)
+rmse_snaive_weekly
 
@@
 
-#### Step 5b: Extract values for time series and plot the series - Monthly
+#### Step 9: Forecast using SARIMA - Weekly
 
-```{r, tidy=TRUE}
-# Create monthly data
-pm2_5_monthly <- apply.monthly(pm2_5_xts, mean)
-
-pm2_5_monthly_ts <- ts(pm2_5_monthly["19990131/20230811"], start = c(1999,1), frequency = 12)
-plot(pm2_5_monthly_ts)
-
-kpss.test(pm2_5_monthly_ts)
-
-adf.test(pm2_5_monthly_ts)
-
-acf(pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
-pacf(pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
-
-```
-
-#### Step 6b: Split data into train and test - Monthly
-
-```{r, tidy=TRUE}
-# Split the data into a training and test dataset
-train_monthly <- window(pm2_5_monthly_ts, start = c(1999,1), end=c(2020,12))
-test_monthly <- window(pm2_5_monthly_ts, start = c(2021,1))
-```
-
-#### Step 7b: Decompose the time series plot - Monthly
-
-```{r, tidy=TRUE}
-
-# Looking at spectral analysis
-periodogram(train_monthly, log = "no", plot=TRUE, 
-            ylab = "Periodogram",
-            xlab = "Frequency",
-            lwd=2, xlim = c(0, 0.2))
-
-# There are two trends: 
-#  1) A typical yearly (12 months) seasonal trend
-#  2) A trend that is repeating every 8-9 years (100 months)
-#  3) A typical half yearly (6 months) seasonal trend
-
-# Overall, there is no mixed effect that normal seasonality model cannot capture.
-
-# Decompose the series
-plot(decompose(train_monthly, type="multiplicative"))
-
-# Important Inferences
-# 1) The PM2.5 AQI has been decreasing overall though there are rise every now and then.
-#    However the trend is going down with time.
-# 2) Winter months have a strong seasonal effect with Nov and Dec being the peak months
-#    usually which likely could be due to cold temperatures causing pollutant to
-#    not escape the lower atmosphere.
-#    Link: https://www.accuweather.com/en/health-wellness/why-air-pollution-is-worse-in-winter/689434#:~:text=Cold%20air%20is%20denser%20and%20moves%20slower%20than%20warm%20air,rate%20than%20during%20the%20summer.
-# 3) We can see a seasonal cycle of 12 months where the mean value of each month starts 
-#    with a increasing trend in the beginning of the year and drops down towards 
-#    the end of the year. We can see a seasonal effect with a cycle of 12 months.
-
-# Understand the seasonality and remove it to see the trend
-train_monthly_diff <- diff(train_monthly, lag = 12)
-autoplot(train_monthly_diff, ylab="Train Seasonal Differencing (Monthly)")
-
-# Let's look if the series is stationary?
-kpss.test(train_monthly_diff)
-# The reported p-value is 0.1, which is > 0.05, and would suggest that we fail 
-# to reject the null hypothesis of level stationarity (conclusion: stationary)
-
-adf.test(train_monthly_diff)
-# The reported p-value of 0.01 (<0.05) so we do reject the null hypothesis that 
-# the time series is non-stationary (conclusion: stationary)
-
-acf(train_monthly_diff, main = "ACF of Seasonal Differencing Time Series")
-pacf(train_monthly_diff, main = "PACF of Seasonal Differencing Time Series")
-
-```
-
-#### Step 8b: Benchmark using snaive model - Monthly
-
-```{r, tidy=TRUE}
-
-# Forecast the seasonal naïve for (2021-01 to 2023-07)
-forecast_snaive_monthly <- snaive(train_monthly, h=31)
-# Plot the forecasts for snaive model
-plot(forecast_snaive_monthly, main = "PM 2.5 AQI Forecast - SNaive",
-         xlab = "Week", ylab = "PM 2.5 AQI")
-lines(test_monthly)
-
-# Compare the forecast with the actual test data by calculating MAPE and MSE
-# Mean Absolute Percentage Error (MAPE)
-mape_snaive <- mean(abs((test_monthly - forecast_snaive_monthly$mean)/test_monthly))
-mape_snaive
-
-# Mean Squared Error (MSE)
-mse_snaive <- mean((test_monthly - forecast_snaive_monthly$mean)^2)
-mse_snaive
-
-# Evaluate the residuals
-checkresiduals(forecast_snaive_monthly)
-
-```
-
-#### Step 9a: TBATS model - Weekly
-
-```{r, tidy=TRUE}
-vec <- as.vector(train_weekly)
-
-univariate_ts <- ts(vec, start=1999, frequency=52.18)
-
-stl_decomposition <- stl(univariate_ts, s.window="periodic")
-plot(stl_decomposition)
-
-h <- length(test_weekly)
-
-model_tbats <- tbats(train_weekly)
-forecast_tbats <- forecast(model_tbats, h=h)
-plot(forecast_tbats)
-
-checkresiduals(model_tbats)
-#Ljung-Box test .05179
-
-plot(pm2_5_weekly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_weekly_ts), max(pm2_5_weekly_ts)), main="Train, Test, and Forecasted Data - Weekly", xlab="Year", ylab="PM2.5 AQI Value")
-lines(test_weekly, col="red") # Test data in red
-lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
-legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
-
-mae <- mean(abs(test_weekly - forecast_tbats$mean))
-print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
-#Mean Absolute Error (MAE): 9.31
-
-comparison_df <- data.frame(
-  Date = time(test_weekly),
-  Actual = as.vector(test_weekly),
-  Forecasted = as.vector(forecast_tbats$mean)
-)
-print(comparison_df)
-
-```
-
-#### Step 9b: TBATS model - Monthly
-
-```{r, tidy=TRUE}
-
-vec <- as.vector(train_monthly)
-
-univariate_ts <- ts(vec, start=1999, frequency=52.18)
-
-stl_decomposition <- stl(univariate_ts, s.window="periodic")
-plot(stl_decomposition)
-
-h <- length(test_monthly)
-
-model_tbats <- tbats(train_monthly)
-forecast_tbats <- forecast(model_tbats, h=h)
-plot(forecast_tbats)
-
-checkresiduals(model_tbats)
-#Ljung-Box test .07956
-
-plot(pm2_5_monthly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_ts), max(pm2_5_monthly_ts)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
-lines(test_monthly, col="red") # Test data in red
-lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
-legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
-
-mae <- mean(abs(test_monthly - forecast_tbats$mean))
-print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
-#Mean Absolute Error (MAE): 6.61
-
-comparison_df <- data.frame(
-  Date = time(test_monthly),
-  Actual = as.vector(test_monthly),
-  Forecasted = as.vector(forecast_tbats$mean)
-)
-print(comparison_df)
-
-```
-
-#### Step 10a: ARIMA model using auto.arima() - Weekly
 ```{r}
@@
 
-```{r}
+```{r, tidy=TRUE}
+# Forecast the TBATS for (2021-01 to 2023-07)
+h <- length(test_weekly)
 
-kpss_test_result <- kpss.test(train_weekly_diff)
-adf_test_result <- adf.test(train_weekly_diff)
+model_sarima_weekly <- auto.arima(train_weekly, 
+                                  seasonal = TRUE, 
+                                  stepwise = FALSE, #stepwise = TRUE,
+                                  trace = FALSE,
+                                  #m=52,
+                                  #d=None,
+                                  #D=1, # order of seasonal differencing
+                                  lambda="auto",
+                                  start.p = 1, max.p = 3,
+                                  start.q = 2, max.q = 4)
 
-cat("KPSS: ", round(kpss_test_result$p.value, 2), "\n")
-cat("ADF: ", round(adf_test_result$p.value, 2), "\n")
+forecast_sarima_weekly <- forecast(model_sarima_weekly, h)
+plot(forecast_sarima_weekly)
 
-# Fit ARIMA model using auto.arima()
-arima_model_weekly <- auto.arima(train_weekly, seasonal = TRUE, stepwise = FALSE,
-                                 trace = FALSE,
-                                 start.p = 1, max.p = 4,
-                                 start.q = 2, max.q = 5)
+# Plot the forecasts for TBATS model
+plot(pm2_5_weekly_ts, 
+     xlim=c(1999, 2023), 
+     ylim=c(min(pm2_5_weekly_ts), 
+            max(pm2_5_weekly_ts)), 
+     main="Train, Test, and Forecasted Data - Weekly", 
+     xlab="Year", 
+     ylab="PM2.5 AQI Value")
+lines(test_weekly, col="red") # Test data in red
+lines(forecast_sarima_weekly$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", 
+       legend=c("Train", "Test", "Forecast"), 
+       fill=c("black", "red", "blue"))
 
-# auto.arima(pm2_5_weekly_diff, lambda = "auto", seasonal=TRUE, )
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Symmetric Mean Absolute Percentage Error (sMAPE)
+smape_sarima_weekly <- smape(test_weekly, forecast_sarima_weekly$mean)
+smape_sarima_weekly
 
-# Forecast using ARIMA model
-h <- length(test_weekly)
-forecast_arima_weekly <- forecast(arima_model_weekly, h = h)
+# Mean Absolute Error (MAE)
+mae_sarima_weekly <- mean(abs((test_weekly - forecast_sarima_weekly$mean)))
+mae_sarima_weekly
 
-plot(forecast_arima_weekly, main = "PM 2.5 AQI Forecast - sARIMA",
-         xlab = "Week", ylab = "PM 2.5 AQI")
-lines(test_weekly)
+# Root Mean Squared Error (RMSE)
+rmse_sarima_weekly <- rmse(test_weekly, forecast_sarima_weekly$mean)
+rmse_sarima_weekly
 
-```
-
-```{r}
-
-# Check the residuals of the ARIMA model
-checkresiduals(arima_model_weekly)
-
-# Calculate MAE
-mae <- mean(abs(forecast_arima_weekly$mean - test_weekly))
-# Calculate MAPE
-mape <- mean(abs((forecast_arima_weekly$mean - test_weekly) / test_weekly)) * 100
-
-# Get AICc and BIC values from the ARIMA model
-aic <- arima_model_weekly$aic
-bic <- arima_model_weekly$bic
-
-# Corrected AICc calculation
-n <- length(train_weekly)
-k <- length(arima_model_weekly$coef)
-aic_corrected <- aic + 2 * (k + 1) * (k + 2) / (n - k - 2)
-
-# Print the results
-cat("MAE: ", round(mae, 2), "\n")
-cat("MAPE: ", round(mape, 2), "%\n")
-cat("AICc: ", round(aic_corrected, 2), "\n")
-cat("BIC: ", round(bic, 2), "\n")
-
-# Create a time index for the forecasted values
-forecast_time_index <- seq(from = end(test_weekly) + 1 / 52, 
-                           length.out = length(forecasted_values), by = 1 / 52)
-
-# Plot historical data, actual test, and forecast
-plot(train_weekly, xlim=c(1999, 2023), ylim=c(min(train_weekly), max(train_weekly)), main="Train, Test, and Forecasted Data - Weekly", xlab="Year", ylab="PM2.5 AQI Value")
-lines(test_weekly, col="red") # Test data in red
-lines(forecast_arima_weekly$mean, col="blue", type="o") # Forecasted data in blue
-legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))
+# Evaluate the residuals
+checkresiduals(forecast_sarima_weekly)
 
@@
   Actual = as.vector(test_weekly),
-  Forecasted = as.vector(forecast_arima_weekly$mean)
+  Forecasted = as.vector(forecast_sarima_weekly$mean)
 )
 print(comparison_df)
-```
-```{r}
-summary(arima_model_weekly)
+
 ```
 
-```{r}
-# Assuming your residuals are stored in a variable called "residuals"
-shapiro_test_result <- shapiro.test(arima_model_weekly$residuals)
+#### Step 10: Forecast using TBATS - Weekly
 
-# Print the test result
-print(shapiro_test_result)
+```{r, tidy=TRUE}
+# Forecast the TBATS for (2021-01 to 2023-07)
+h <- length(test_weekly)
 
-# Check the p-value
-if (shapiro_test_result$p.value < 0.05) {
-  cat("The residuals are not normally distributed (p-value:", shapiro_test_result$p.value, ")\n")
-} else {
-  cat("The residuals appear to be normally distributed (p-value:", shapiro_test_result$p.value, ")\n")
-}
+vec <- as.vector(train_weekly)
+univariate_ts <- ts(vec, start=1999, frequency=52.18)
+plot(stl(univariate_ts, s.window="periodic"))
+
+model_tbats_weekly <- tbats(train_weekly,
+                            use.box.cox = TRUE)
+forecast_tbats_weekly <- forecast(model_tbats_weekly, h)
+plot(forecast_tbats_weekly)
+
+# Plot the forecasts for TBATS model
+plot(pm2_5_weekly_ts, 
+     xlim=c(1999, 2023), 
+     ylim=c(min(pm2_5_weekly_ts), 
+            max(pm2_5_weekly_ts)), 
+     main="Train, Test, and Forecasted Data - Weekly", 
+     xlab="Year", 
+     ylab="PM2.5 AQI Value")
+lines(test_weekly, col="red") # Test data in red
+lines(forecast_tbats_weekly$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", 
+       legend=c("Train", "Test", "Forecast"), 
+       fill=c("black", "red", "blue"))
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Symmetric Mean Absolute Percentage Error (sMAPE)
+smape_tbats_weekly <- smape(test_weekly, forecast_tbats_weekly$mean)
+smape_tbats_weekly
+
+# Mean Absolute Error (MAE)
+mae_tbats_weekly <- mean(abs((test_weekly - forecast_tbats_weekly$mean)))
+mae_tbats_weekly
+
+# Root Mean Squared Error (RMSE)
+rmse_tbats_weekly <- rmse(test_weekly, forecast_tbats_weekly$mean)
+rmse_tbats_weekly
+
+# Evaluate the residuals
+checkresiduals(forecast_tbats_weekly)
+
+comparison_df <- data.frame(
+  Date = time(test_weekly),
+  Actual = as.vector(test_weekly),
+  Forecasted = as.vector(forecast_tbats_weekly$mean)
+)
+print(comparison_df)
+
 ```
+
+#### Step 11: Forecast using Neural Network - Weekly
+
+```{r, tidy=TRUE}
+h <- length(test_weekly)
+model_nn_weekly <- nnetar(train_weekly,
+                           size=10,
+                           decay=0.5,
+                           repeats=1000,
+                           lambda="auto")
+summary(model_nn_weekly)
+forecast_nn_weekly <- forecast(model_nn_weekly, PI=FALSE, h)
+plot(forecast_nn_weekly)
+
+# Plot the forecasts for NN model
+plot(pm2_5_weekly_ts, 
+     xlim=c(1999, 2023), 
+     ylim=c(min(pm2_5_weekly_ts), 
+            max(pm2_5_weekly_ts)), 
+     main="Train, Test, and Forecasted Data - Weekly", 
+     xlab="Year", 
+     ylab="PM2.5 AQI Value")
+lines(test_weekly, col="red") # Test data in red
+lines(forecast_nn_weekly$mean, col="blue", type="o") # Forecasted data in blue
+legend("topleft", 
+       legend=c("Train", "Test", "Forecast"), 
+       fill=c("black", "red", "blue"))
+
+# Compare the forecast with the actual test data by calculating MAPE and MSE
+# Symmetric Mean Absolute Percentage Error (sMAPE)
+smape_nn_weekly <- smape(test_weekly, forecast_nn_weekly$mean)
+smape_nn_weekly
+
+# Mean Absolute Error (MAE)
+mae_nn_weekly <- mean(abs((test_weekly - forecast_nn_weekly$mean)))
+mae_nn_weekly
+
+# Root Mean Squared Error (RMSE)
+rmse_nn_weekly <- rmse(test_weekly, forecast_nn_weekly$mean)
+rmse_nn_weekly
+
+# Evaluate the residuals
+checkresiduals(forecast_nn_weekly)
+
+comparison_df <- data.frame(
+  Date = time(test_weekly),
+  Actual = as.vector(test_weekly),
+  Forecasted = as.vector(forecast_nn_weekly$mean)
+)
+print(comparison_df)
+
+```

### 2023-08-16 MSCA_31006_Time_Series_Final_Combined.Rmd
Update MSCA_31006_Time_Series_Final_Combined.Rmd

@@
                                   seasonal = TRUE, 
-                                  stepwise = FALSE, #stepwise = TRUE,
-                                  trace = FALSE,
-                                  #m=52,
-                                  #d=None,
-                                  #D=1, # order of seasonal differencing
                                   lambda="auto",
-                                  start.p = 1, max.p = 3,
-                                  start.q = 2, max.q = 4)
+                                  start.p = 1, 
+                                  max.p = 3,
+                                  start.q = 2, 
+                                  max.q = 4)
 
